The {\em Diff} MILP model and PCA for \\ {\em real-time} DNN robustness certification

### 2025-11-20 CAV 2026/global.tex
exp + global

@@

    To encode global robustness, \cite{vhagar} considers
	variable $x_j,\hat{x}_j$ and $x'_j,\hat{x}'_j$ for nodes of the image $I$ 
	and of its perturbation $I'$, 
	using the classical encoding (\ref{eq11}) \cite{MILP} 
	to compute $\hat{x}_i, \hat{x}'_i$ from $x_i, x'_i$
	for each $i$.
    Further:
    \begin{enumerate}
		\item For each input nodes $x_j,x'_j$ of $I,I'$, add constraints
        $|x_j-x'_j| \leq \varepsilon$ 
        ($L_\infty$-perturbation - see later for $L_1$).
		\item The optimization objective is to maximize $x_C - x_D +x'_D - x'_C$ 
		to compute $\beta^\varepsilon_{C,D}=\beta^\varepsilon_{D,C}$, 
		for $C \neq D$  decision classes, as proposed in \cite{vhagar}.
	\end{enumerate}

>>>>

	\subsection{Optimization objective for real-time verification}

As explained in the introduction, the objective we want to maximize is 
$\beta^{\text{perturb}}_{C,D} = x_C - x_D +x'_D - x'_C$, 
for $x_C,x_D$ the value associated with the output neurons $C,D$ from input image $I$, and $x'_C,x'_D$ from perturbation $I'$. These bounds 
$\beta^{\text{perturb}}_{C,D}$ will allow us to verify in real-time 
whether a new incoming image can be certified to be robust for the perturbation.

Different bounds were proposed in \cite{vhagar}, for an evaluation purpose rather than to verify images in real-time.

TODO: DEF OF BOUNDS IN VHAGAR. CALL BOUNDS $\alpha_C,D$, to compute the perturbation necessary to switch an image classified as $C$ into an image classified as $D$.
Notice that $\alpha_C,D$ is not symetrical, while 
$\beta^{\text{perturb}}_{C,D}$ is.

@@

    To encode global robustness in MILP, the classical way, used (with variants) by \cite{vhagar,ITNE}, is to consider variables $x_j,\hat{x}_j$ and 
	$x'_j,\hat{x}'_j$, encoding the value of neuron $j$ obtained from the image $I$ and from its perturbation $I'$, respectively. Further:
	\begin{enumerate}
		\item For each input neuron $j,j$ of $I,I'$, there is constraint
        $|x_j-x'_j| \leq \varepsilon$ to encode the $L_\infty$-perturbation 
		(see later how we encode $L_1$ perturbations).
		\item For each neuron $j$, we have the Re ReLU constraints 
		from the classical encoding (\ref{eq11}) \cite{MILP}
		for both $\hat{x}_j$ and $\hat{x'}_j$ independently.
	\end{enumerate}

### 2025-11-20 CAV 2026/introduction.tex
exp + global

	\item We develop a novel {\em diff MILP encoding} for the global robustness problem, %called the {\em "2v" model}, 
	where the variables are the values of the perturbed neurons, as well as the difference between the original and the perturbed neuron values (called the {\em diff variables}, introduced in \cite{diff}). We study and encode how the {\em diff variables} evolve after passing through a ReLU (Prop.~\ref{Prop2}). Compared with the {\em classical MILP model} \cite{MILP} employed in \cite{vhagar,lipshitz,ITNE}, which considers the input and the perturbation but not their difference ({\em diff variable}), we keep the same number of 2 binary variables per ReLU. 
	%the results are more accurate for the same runtime.
	%the linear relaxation is much more accurate, as each {\em diff variable} can be bounded after a ReLU as a function of the value of the {\em diff variables} before the ReLU, whereas the linear relaxations of the classical model is extremely inaccurate, as the variables are independent of each other. This was observed in \cite{lipshitz,ITNE}, and constraints encoding
	%a part of the linear relaxation of our "2v" model were added explicitly. 
	%Experimentally, 
	However, our {\em diff MILP model} is more efficient, one reason being 
	the accuracy of its linear relaxation.
	%The number of variables a priori doubles compared with local robustness, from each neuron value in the perturbed image to each neuron value in the perturbed image {\em and} in the original image, as the original image is no more fixed. 
	%Recall that the worst case complexity of MILP is exponential in the number of binary variables \cite{DivideAndSlide}. 
	%A straightforward MILP model would be to use the  for each of these variables, as in \cite{vhagar,lipshitz}. The main issue with the classical model is that its linear relaxations is extremely inaccurate, as the variables are independent of each other. Instead, we develop another exact MILP model, 
	\item Further from the {\em diff MILP model}, which is exact, 
	we develop two abstract MILP models, which are more efficient but also possibly less accurate.
	Namely, the "1b+2" model replaces the 2 binary variables by 1 binary (for the diff variable) and 2 linear variables (linear relaxation for the perturbation); while the "1b" model has a unique binary variable per ReLU, only considering the {\em diff variables}. 
	%\item We adapt the Solution Aware Scoring from \cite{ATVA25} to our novel MILP models, in order to select the most important ReLUs to be treated using complex binary variables, while less important variables are treated using linear relaxation. The chosen number of binary variables depends upon the complexity of the DNN as well as the targeted runtime.

>>>>

	\item We develop a novel {\em diff MILP encoding} for the global robustness problem, %called the {\em "2v" model}, 
	where the variables are the values of the perturbed neurons, as well as the difference between the original and the perturbed neuron values (called the {\em diff variables}, introduced in \cite{diff}). We study and encode how the {\em diff variables} evolve after passing through a ReLU (Prop.~\ref{Prop2}). Compared with the {\em classical MILP model} \cite{MILP} employed in \cite{vhagar,lipshitz,ITNE}, which considers the input and the perturbation but not their difference ({\em diff variable}), we keep the same number of 2 binary variables per ReLU. 
	%the results are more accurate for the same runtime.
	%the linear relaxation is much more accurate, as each {\em diff variable} can be bounded after a ReLU as a function of the value of the {\em diff variables} before the ReLU, whereas the linear relaxations of the classical model is extremely inaccurate, as the variables are independent of each other. This was observed in \cite{lipshitz,ITNE}, and constraints encoding
	%a part of the linear relaxation of our "2v" model were added explicitly. 
	%Experimentally, 
	However, our {\em diff MILP model} is more efficient, one reason being 
	the accuracy of its linear relaxation.
	%The number of variables a priori doubles compared with local robustness, from each neuron value in the perturbed image to each neuron value in the perturbed image {\em and} in the original image, as the original image is no more fixed. 
	%Recall that the worst case complexity of MILP is exponential in the number of binary variables \cite{DivideAndSlide}. 
	%A straightforward MILP model would be to use the  for each of these variables, as in \cite{vhagar,lipshitz}. The main issue with the classical model is that its linear relaxations is extremely inaccurate, as the variables are independent of each other. Instead, we develop another exact MILP model, 
	\item Further from the {\em diff MILP model}, which is exact, 
	we develop two abstract MILP models, which are more efficient but also possibly less accurate. Namely, the "2b+1" model, decoupling one binary variable into 2 variables: 1 binary and the other using linear relaxation; while the "1b" model has a unique binary variable per ReLU, only considering the {\em diff variables}.
	%\item We adapt the Solution Aware Scoring from \cite{ATVA25} to our novel MILP models, in order to select the most important ReLUs to be treated using complex binary variables, while less important variables are treated using linear relaxation. The chosen number of binary variables depends upon the complexity of the DNN as well as the targeted runtime.
